Relationships in the extremes and their influence on competition and coexistence: Supporting information

### 2022-02-10 paper/SupMat.Rnw
finish sm section 'Details of diatom model setup'

%***DAN: Jasmin, please start on this section when you get a chance. Here are some topics that need
%to be covered:
Estimations of $Q_i(t)$, $V_i(t)$, and $K_i(t)$, temperature dependence functions for 
cell quota, reproduction rate, and half saturation, respectively, were used to 
paramaterize the competing diatom model outline in the paper. 
Values of and subsequent function estimations are plotted in Figure \ref{fig:VQK}. 
\cite{Descamps_2005} obtained estimated values for $Q_i$, $V_i$, and $K_i$ from 
single species nine-day batch experiments at four different temperatures ($6^\circ$C, 
$12^\circ$C, $18^\circ$C, $24^\circ$C) (see Table 1 of \cite{Descamps_2005}) and further
computed linear interpolations to
functions $Q_i(t)$, $V_i(t)$, and $K_i(t)$. These values and functions are 
indicated in Figure \ref{fig:VKQ} by open circles and dashed lines. 
\cite{Ellne_2016, Ellner_2019} chose to slightly modify some values and functions 
(to better align model with competition experiments), 
%why did they make K_i's constant and do we need to adjust so that it
%is constant over the range of temperatures considered 
indicated in Figure \ref{fig:VKQ} by closed circles and solid lines. 
These modifications do not result in meaningful differences when evaluating 
coexistence at the parameter values used by \cite{Descamps_2005}.
We chose to implement the modified functiones in our methods to quantify and 
decompose coexistence.

@@

Our decomposition is defined by $E(t)=\theta(t)$, 
$C_i(t)=(K_i(t)+S(t))/S(t)$ and $r_i(E_i,C_i)=\frac{V_i(E)}{C_i} -D$. 
We note that our definition and selection of our "features" of interest are
options among many and features need not be limited to $E$ and $C$ \citep{Ellner_2019}. 
For example \cite{Ellner_2019} decomposes coexistence with respect to fluctuation in
inflow silicate concentration, $S$, and temperature, and \cite{Ellner_2016} decomposes
coexistence using classical $E$ and $C$ drivers, but define $E=V(t)$. 
Like \cite{Ellner_2016}, we set $C_i(t)=(K_i(t)+S(t))/S(t)$ to reflect 
potential versus actual growth of the diatom populations. 
Choice of drivers and their definitions are up to the investigator,
but should refelct real variables that influence population growth.

%3) You need to outline how the quantities E_i and C_i\i were obtained from simulations. 
To obtain $E_i$ and $C_{i \bs i}$, the model was simulated for a $T=3600$ number of timesteps, setting initial population size of invader i to be zero and initial population of the resident, $r=20$.

>>>>

%***DAN: Jasmin, please start on this section when you get a chance. Here are some topics that need
%to be covered:
Estimations of $Q_i(t)$, $V_i(t)$, and $K_i(t)$, temperature dependence functions for 
cell quota, reproduction rate, and half saturation, respectively, were used to 
paramaterize the competing diatom model outlined in the paper. 
Values of and subsequent function estimations are plotted in Figure \ref{fig:VQK}. 
\cite{Descamps_2005} obtained estimated values for $Q_i$, $V_i$, and $K_i$ from 
single species nine-day batch experiments at four different temperatures ($6^\circ$C, 
$12^\circ$C, $18^\circ$C, $24^\circ$C) (see Table 1 of \cite{Descamps_2005}) and further
computed linear interpolations to
functions $Q_i(t)$, $V_i(t)$, and $K_i(t)$. These values and functions are 
indicated in Figure \ref{fig:VKQ} by open circles and dashed lines. 
\cite{Ellne_2016, Ellner_2019} used slightly modified values and functions,
indicated in \ref{fig:VQK} by filled circles and solid lines. 
Our investigation follows that of \cite{Ellne_2016, Ellner_2019} and 
uses the modified functions of  $Q_i(t)$, $V_i(t)$, and $K_i(t)$ to analyze coexistence.

@@

Our decomposition is defined by $E(t)=\theta(t)$, 
$C_{i \bs i}(t)=(K_i(t)+S(t))/S(t)$ and $r_i(E_i,C_{i \bs i})=\frac{V_i(E)}{C_{i \bs i}} -D$. 
We note that our definition and selection of our "features" of interest are
options among many and features need not be limited to $E$ and $C$ \citep{Ellner_2019}. 
For example \cite{Ellner_2019} decomposes coexistence with respect to fluctuation in
inflow silicate concentration, $S$, and temperature, and \cite{Ellner_2016} decomposes
coexistence using classical $E$ and $C$ drivers, but define $E=V(t)$. 
Like \cite{Ellner_2016}, we set $C_{i \bs i}(t)=(K_i(t)+S(t))/S(t)$ to reflect 
potential versus actual growth of the diatom populations. 
Choice of drivers and their definitions are up to the investigator,
but should refelct real variables that influence population growth.

%3) You need to outline how the quantities E_i and C_i\i were obtained from simulations. 
To obtain $E_i$ and $C_{i \bs i}$, the competiton model was simulated for a $T=3000$ number of timesteps to obtain $S$, the extracellular silicate concentration. The initial population size of invaderwas set at zero and initial population of the resident was set at 20. Environmental temperature fluctuations were defined by the paramters $a$, $P$, and $\theta_0$ for $\theta(t)=asin(2\pi t/P)+\theta_0$. Dilution was set to $D=0.09$ and inflow silicate concentration was set at $S_0=35$. A burn in time of $\frac{2}{3}T$ was cut from the resulting $S$ and $E_i(t)=\theta(t)$. Competitive pressure was calculated using $C_{i \bs i}(t)=(K_i(t)+S(t))/S(t)$.
